---
title: "p8131_hw4_ps3194"
author: "Pangsibo Shen"
date: "2/19/2021"
output: html_document
---

```{r}
library(dplyr)
library(nnet)
library(MASS)
```

#### 1)

```{r table}
contact_level = c(rep("low", 3), rep("high",3))
house_type = rep(c("Tower_block", "Apartment", "House"), 2)
sat_low = c(65,130,67,34,141,130)
sat_med = c(54,76,48,47,116,105)
sat_high = c(100,111,62,100,191,104)

housing_df =
  tibble(contact_level,house_type,sat_low,sat_med,sat_high) %>%
  mutate(
    contact_level = as.factor(contact_level),
    house_type = as.factor(house_type)
         )
  

#Table of percentages showing the pair-wise associations between satisfaction and contact level
housing_df %>% dplyr::select(contact_level,sat_low,sat_med, sat_high) %>%
  tidyr::pivot_longer(cols = starts_with("sat"),
               names_to = "sat_level",
               names_prefix = "sat_",
               values_to = "freq") %>%
  group_by(contact_level,sat_level) %>%
  summarise(
    total = sum(freq)
  ) %>%
  tidyr::pivot_wider(
    names_from = sat_level,
    values_from = total,
    names_prefix = "sat_"
  ) %>%
  mutate(total_freq = sat_high + sat_low + sat_med,
         sat_low_pct = (sat_low/total_freq)*100,
         sat_med_pct = (sat_med/total_freq)*100,
         sat_high_pct = (sat_high/total_freq)*100
         ) %>%
  dplyr::select(contact_level,total_freq,sat_low_pct,sat_med_pct,sat_high_pct) %>%
  knitr::kable()

#Table of percentages showing the pair-wise associations between satisfaction and house type
housing_df %>% dplyr::select(house_type,sat_low,sat_med, sat_high) %>%
  tidyr::pivot_longer(cols = starts_with("sat"),
               names_to = "sat_level",
               names_prefix = "sat_",
               values_to = "freq") %>%
  group_by(house_type,sat_level) %>%
  summarise(
    total = sum(freq)
  ) %>%
  tidyr::pivot_wider(
    names_from = sat_level,
    values_from = total,
    names_prefix = "sat_"
  ) %>%
  mutate(total_freq = sat_high + sat_low + sat_med,
         sat_low_pct = (sat_low/total_freq)*100,
         sat_med_pct = (sat_med/total_freq)*100,
         sat_high_pct = (sat_high/total_freq)*100
         ) %>%
  dplyr::select(house_type,total_freq,sat_low_pct,sat_med_pct,sat_high_pct) %>%
  knitr::kable()

```
From the table of percentages showing the pair-wise associations between satisfaction and contact level, we learned that among residents with high contact level, residents appear to have higher probability to have high satisfaction and among residents with low contact level, residents appear to have the lowest probability to have medium satisfaction. 

From table of percentages showing the pair-wise associations between satisfaction and house type, we learned that among residents living in the apartments, residents appear to have higher probability to have high satisfaction;among residents living in the houses, residents appear to have the highest probability to have low satisfaction; among residents living in the Tower block, residents appear to have the highest probability to have high satisfaction.

#### 2)

```{r nomial logistic regression}
housing_nomial = multinom(cbind(sat_low, sat_med, sat_high)~contact_level+house_type, data = housing_df)
summary(housing_nomial)

# exponentiate the coefficients for interpretations
exp(summary(housing_nomial)$coefficients)
```


```{r confidence interval}
#odds ratio with 95% confidence interval
coef_df = 
  summary(housing_nomial)$coefficients

std_err_df=
  summary(housing_nomial)$standard.errors %>%
  as_tibble() %>%
  dplyr::select(-1) %>%
  tidyr::pivot_longer(cols = c(contact_levellow, house_typeHouse, house_typeTower_block),
               names_to = "covariates",
               values_to = "std_err") %>%
  dplyr::select(std_err)

or_ci =
  tibble(
    term = c("med_sat_contactLow","med_sat_typehouse","med_sat_typetowerblock","high_sat_contactLow","high_sat_typehouse","high_sat_typetowerblock"),
    point_est = c(coef_df[3],coef_df[5],coef_df[7],coef_df[4],coef_df[6],coef_df[8]),
    std_err_df
  ) %>%
  mutate(
    exp_point_est = exp(point_est),
    lower_exp_ci = exp(point_est + qnorm(0.025)*std_err),
    higher_exp_ci = exp(point_est - qnorm(0.025)*std_err)
  )
or_ci %>%
  knitr::kable()
```

The table with exponential point estimates and confidence intervals for odds ratios is shown above. 


```{r goodness of fit}
# goodness of fit
pihat=predict(housing_nomial,type='probs') 
m=rowSums(housing_df[,3:5])
res_pearson=(housing_df[,3:5]-pihat*m)/sqrt(pihat*m) # pearson residuals 

G_stat=sum(res_pearson^2) # Generalized Pearson Chisq Stat
G_stat

pval=1-pchisq(G_stat,df = (6-4)*(3-1)) 
pval# fit is good
```
Since the p-value for Generalized Pearson Chisq Statistic is greater than 0.05, we can conclude that our model is a good fit.

#### 3）

```{r ordinal logistic regression}
#create a dataframe in long format which can be used in proportional odds model
housing_df_long = 
  housing_df %>%
  dplyr::select(contact_level,house_type,sat_low,sat_med, sat_high) %>%
  tidyr::pivot_longer(cols = starts_with("sat"),
               names_to = "sat_level",
               names_prefix = "sat_",
               values_to = "freq") %>%
  mutate(sat_level = factor(sat_level,levels = c("low", "med", "high")))

# fit proportional odds model
housing_prop = polr(sat_level~contact_level+house_type,data = housing_df_long,weights = freq)
summary(housing_prop)

# exponentiate the coefficients for interpretations
exp(0.2524) #contact_levellow
exp(0.2353) #house_typeHouse       
exp(-0.5010) #house_typeTower_block  
```
The fitted model told us that: the odds ratio of low satisfaction vs medium/high satisfaction between residents with low contact level and residents with high contact level is `r exp(0.2524)`, while holding other variable fixed;the odds ratio of low satisfaction vs medium/high satisfaction between residents living in house and residents not living in house is `r exp(0.2353)`, while holding other variable fixed;the odds ratio of low satisfaction vs medium/high satisfaction between living in tower block and residents not living in tower block is `r exp(-0.5010)`, while holding other variable fixed.

#### 4）

```{r pearson residuals}
pi_prop = predict(housing_prop,housing_df, type = "p")
m_prop = rowSums(housing_df[,3:5])
res_pearson_prop = (housing_df[,3:5]-pi_prop*m_prop)/sqrt(pi_prop*m_prop)
res_pearson_prop %>%
  mutate(
    contact_level = c(rep("low", 3), rep("high",3)),
    house_type = rep(c("Tower_block", "Apartment", "House"), 2)
  ) %>%
  knitr::kable()
```

The largest discrepancy between the observed frequencies and expected frequencies estimated from the model happened when the contact level is high, housing type is house, and satisfactory level is high. It has the largest pearson residual of -1.48.

